Overview \& Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-22 WSSSPE1_report/WSSSPE1_summary_report.tex
More minor changes.

\title[]{Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)}
%\note{\scriptsize Authors, please add/check middle initials and affiliations}

@@

Software is developed for many different purposes, and the requirements can
vary significantly depending on the intended audience. Most end-users make use
of either a graphical user interface of some kind, or a command line that may
offer input and output formats for running the code and analyzing its output(s).
When discussing backward compatibility it is these various interfaces that are
discussed. For software that builds on other software frameworks it is the
API that is most important, and this can
encompass issues such as the source and binary interfaces to the software
libraries developed---with each potentially having a high maintenance cost if
they are to remain compatible over many years. When using command-line programs
it is generally the command-line switches as well as the input/output formats
that could incur significant costs if they are changed.

@@

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center (i.e., a place to `take care' of software
when the original developers have stopped doing so).
The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g., Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

The role of the growing field of team science with software development  was discussed. The science of team science
deals with understanding and improving collaborative and team-based scientific research, and such issues as virtual organizations,
and tool development cut across software development communities. Further, how should these skills and best practices then be
introduced to students?

@@

Additional activities that are being planned include
two additional WSSSPE workshops at the 2014 SciPy and SC14 conferences.
The SciPy workshop (WSSSPE1.1) will focus on how some software projects
from the SciPy community have dealt with software sustainability issues, both
successfully and unsuccessfully, while the SC14 workshop (WSSSPE2) will
be more general, and will likely focus on determining and publicizing specific
activities that the overall scientific software community can take to move forward.
In addition, there will be a minisymposium on ``Reliable Computational Science''
at the 2014 Society for Industrial and Applied Mathematics (SIAM)
Annual Meeting (SIAM AN14, \url{http://meetings.siam.org})
to further explore some of the key issues raised in
WSSSPE.

@@

Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation.

>>>>

\title[]{Overview \& Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)}
%\note{\scriptsize Authors, please add/check middle initials and affiliations}

@@

Software is developed for many different purposes, and the requirements can
vary significantly depending on the intended audience. Most end-users make use
of either a graphical user interface of some kind, or a command line that may
offer input and output formats for running the code and analyzing its output(s).
When discussing backward compatibility it is these various interfaces that are
discussed. For software that builds on other software frameworks it is the
APIs that are most important, and this can
encompass issues such as the source and binary interfaces to the software
libraries developed---with each potentially having a high maintenance cost if
they are to remain compatible over many years. When using command-line programs
it is generally the command-line switches as well as the input/output formats
that could incur significant costs if they are changed.

@@

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers \note{cite?}
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center (i.e., a place to `take care' of software
when the original developers have stopped doing so).
The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g., Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

The role of the growing field of team science with software development  was discussed. Team science
deals with understanding and improving collaborative and team-based scientific research, and  issues such as virtual organizations,
and tool development across software development communities. Further, how should these skills and best practices then be
introduced to students?

@@

Additional activities that are being planned include
two additional WSSSPE workshops at the 2014 SciPy and SC14 conferences.
The SciPy workshop (WSSSPE1.1) will focus on how some software projects
from the SciPy community have dealt with software sustainability issues, both
successfully and unsuccessfully, while the SC14 workshop (WSSSPE2) will
be more general, and will likely focus on determining and publicizing specific
activities that the overall scientific software community can take to move forward.
In addition, there will be a minisymposium on ``Reliable Computational Science''
at the 2014 Society for Industrial and Applied Mathematics  
Annual Meeting (SIAM AN14, \url{http://meetings.siam.org})
to further explore some of the key issues raised here.

@@

Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author and do not
necessarily reflect the views of the National Science Foundation.
